---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MRCIGVAR

<!-- badges: start -->
<!-- badges: end -->

`MRCIGVAR` provides data-generation, estimation, model-selection, testing, and
impulse-response tools for a family of global and regime-switching
multivariate time-series models. The package covers both standard and
cointegrated specifications, with particular emphasis on multi-regime
cointegrated global VAR systems.

The current refactor standardises the public API around lower snake_case
function names while keeping backward-compatible legacy aliases where needed.

## Supported model families

The package currently includes workflows for:

- `VAR`
- `CIVAR`
- `MRVAR`
- `MRCIVAR`
- `GVAR`
- `CIGVAR`
- `MRGVAR`
- `MRCIGVAR`

Across those families, the package supports:

- synthetic data generation
- parameter estimation
- lag and threshold model selection
- impulse response functions and confidence bands
- regime-migration GIRFs
- constraint and diagnostic routines

## Installation

Install the development version from GitHub with:

``` r
# install.packages("pak")
pak::pak("CurtinIDS/MRCIGVAR")
```

For local development, clone the repository and either load or install the
package locally:

``` r
devtools::load_all()
# or
devtools::install()
```

## Refactored API

The refactored codebase uses a consistent naming scheme:

- `*_data()` for data generators and constructors
- `*_estimate()` for estimation routines
- `*_select_*()` for model-selection helpers
- `irf_*()` and `girf_*()` for response analysis

Examples:

- `cigvar_data()`, `cigvar_estimate()`
- `mrcigvar_data()`, `mrcigvar_estimate()`
- `irf_cigvar_cb()`, `irf_mrcigvar_cb()`

Legacy aliases are still available for backward compatibility, but new code
should use the refactored lowercase function names.

## Example workflow

Below is a minimal CIGVAR workflow using the refactored API.

```{r example-workflow, eval = FALSE}
library(MRCIGVAR)

n <- 4
p <- matrix(0, nrow = n, ncol = 3)
p[, 1] <- 2
p[, 2] <- 2

res_d <- cigvar_data(
  m = 2,
  n = n,
  p = p,
  T = 200,
  type = "const",
  DFYflag = 0
)

res_e <- cigvar_estimate(res_d)
res_e$Summary

sel <- cigvar_select_(res = res_e, L_V = c(3, 3))
sel[which.min(sel[, ncol(sel)]), ]

irf_cb <- irf_cigvar_cb(
  res = res_e,
  nstep = 12,
  comb = NA,
  irf = "gen1",
  runs = 50,
  conf = c(0.05, 0.95)
)

plot_irf(irf_cb, response = c(1, 2), impulse = c(1, 2), ncol = 2)
```

For multi-regime workflows, the package provides parallel families such as
`mrcigvar_data()`, `mrcigvar_estimate()`, `mrcigvar_select_()`, and
`irf_mrcigvar_cb()`.

## Further documentation

For longer-form documentation, see the vignette set:

- `vignettes/mrcigvar-introduction.Rmd`
- `vignettes/core-model-workflows.Rmd`
- `vignettes/mrcigvar-workflow.Rmd`
- `vignettes/irf-and-girf.Rmd`
- `vignettes/refactor-migration-guide.Rmd`

## Project status

This repository is under active refactoring. The main goals of the refactor are:

- consistent public naming
- clearer roxygen documentation
- improved module boundaries in `R/`
- compatibility wrappers for legacy function names

## Documentation and development

- Edit `README.Rmd`, not `README.md`.
- `NAMESPACE` is generated by roxygen2.
- Most files in `R/` are split-module outputs and should be edited carefully.
- The MkDocs site source lives in `docs/` with configuration in `mkdocs.yml`.
- Create the docs environment with
  `conda env create -f environment-docs.yml`.
- Activate it with `conda activate mrcigvar-docs`.
- Preview the docs locally with `mkdocs serve`.

Contributor expectations and project conventions are documented in
`CONTRIBUTING.md`.

## Issues and feedback

Bug reports and feature requests are best submitted through the GitHub issue
tracker:

<https://github.com/CurtinIDS/MRCIGVAR/issues>
