---
title: "Impulse Responses and GIRFs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Impulse Responses and GIRFs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MRCIGVAR)
```

## Purpose

This vignette summarises the impulse-response tooling in `MRCIGVAR`, including:

- standard IRFs
- confidence-band variants
- regime-specific IRFs
- generalized impulse response functions under regime migration

## Naming conventions

The package follows a predictable naming scheme:

- `irf_*()` for impulse response functions
- `irf_*_cb()` for impulse responses with confidence bands
- `girf_*_rm()` for generalized impulse responses with regime migration
- `girf_*_rm_cb()` for GIRFs with confidence bands

Examples:

- `irf_civar_cb()`
- `irf_gvar_cb()`
- `irf_mrcigvar_cb()`
- `girf_mrcigvar_rm()`
- `girf_mrgvar_rm_cb()`

## Common IRF arguments

Across many model families, the main arguments are:

- `res`: the fitted model object
- `nstep`: the number of horizons
- `comb`: optional combined-shock definition
- `irf`: the identification or response type
- `runs`: bootstrap replications
- `conf`: confidence-band tail probabilities

This common interface is one of the main advantages of the refactored API.

## A standard IRF example

```{r cigvar-irf, eval = FALSE}
n <- 4
p <- matrix(0, nrow = n, ncol = 3)
p[, 1] <- 2
p[, 2] <- 2

res_d <- cigvar_data(m = 2, n = n, p = p, T = 200, type = "const")
res_e <- cigvar_estimate(res_d)

irf_cb <- irf_cigvar_cb(
  res = res_e,
  nstep = 12,
  comb = NA,
  irf = "gen1",
  runs = 50,
  conf = c(0.05, 0.95)
)

plot_irf(irf_cb, response = c(1, 2), impulse = c(1, 2), ncol = 2)
```

## Multi-regime IRFs

For multi-regime systems, state selection becomes part of the response
definition.

```{r mrcigvar-irf, eval = FALSE}
irf_cb <- irf_mrcigvar_cb(
  res = res_e,
  nstep = 10,
  comb = NA,
  state = c(2, 1, 1),
  irf = "gen1",
  runs = 20,
  conf = c(0.05, 0.95),
  NT = 1
)
```

In that setting, `state[i]` determines which regime is active for country `i`
during the IRF construction.

## GIRFs under regime migration

For nonlinear response analysis with simulated regime paths, use the GIRF
helpers.

```{r girf-example, eval = FALSE}
girf <- girf_mrcigvar_rm(
  res = res_e,
  shock = c(1, 1, 1, 1, 1, 1),
  R = 100,
  nstep = 10,
  Omega_hist = NA,
  resid_method = "parametric"
)

girf_cb <- girf_mrcigvar_rm_cb(
  res = res_e,
  shock = c(1, 1, 1, 1, 1, 1),
  R = 100,
  nstep = 10,
  Omega_hist = NA,
  resid_method = "parametric",
  conf = c(0.05, 0.95),
  N = 50
)
```

These functions are especially useful when the response depends on the
simulated path of the regime process rather than only on a fixed linear state.

## Identification options

Many IRF helpers accept an `irf` argument with options such as:

- `"gen"`
- `"gen1"`
- `"chol"`
- `"chol1"`
- `"comb1"`

Some model families also support:

- permanent-transitory decompositions
- AB-SVAR style transformations
- explicit structural matrices
- exogenous-shock responses

The lower-level helper `u1()` constructs the shock transformation matrix used by
these routines.

## Plotting

Use `plot_irf()` to visualize IRF arrays returned by many of the high-level
response functions.

```{r plot-example, eval = FALSE}
plot_irf(
  irf_cb,
  Names = NA,
  INames = NA,
  response = c(1, 2),
  impulse = c(1, 2),
  ncol = 2
)
```

## Practical guidance

- Use smaller `runs` values during exploratory work and larger values for final
  analysis.
- Verify output dimensions before downstream use.
- Keep the interpretation of `comb` explicit when constructing concerted or
  combined shocks.
- Prefer state-specific IRFs for linear regime comparisons and GIRFs for
  nonlinear path-dependent analysis.

## Related vignettes

- `mrcigvar-workflow` for the flagship regime-switching workflow
- `core-model-workflows` for standard estimation pipelines
- `mrcigvar-introduction` for the documentation map
